Statistical Grand Slams

### 2023-12-14 Lahman Analysis/Formal Analysis/Draft1.qmd
Finished through beta and eta section

## Baseball Background {#sec-BaseBack}

@@

## Literature Review {#sec-LitReview}

@@

The goal of our model, as described in @eq-model, @eq-priors, and @eq-hyperpriors, was to be able to adaptively predict player $n$'s HR's. That is, we wished to create a model that would predict player $n$'s HR's and would adapt these predictions to be unique to player, $n$, the park played at, $p$, and the year played, $i$. To check the fit of their model, Fellingham and Fisher (2017) selected popular players (such as Hank Aaron and Barry Bonds) as a representative sample of the player population. However, the problem with this appraoch is that Fellingham and Fisher (2017) have selected players who are phenomenal hitters and generally outliers in the data. For example, Barry Bonds currently holds the single season and career HR records. Which means that he has the most HR's hit in one season and the most hit over the course of a career compared to all people who have every played in MLB. The career HR record he holds was previously held by Hank Aaron. Obviously, this sample of players is not representative of the population of players considered by Fellingham and Fisher (2017). So, while it may make the paper more enjoyable to read, it does not provide use with a rigourus understanding of the general fit of this model. We believe that a random sample should have been used to examine the fit of their model since it would give a better sense of overall model fit. Which, is exactly what we did in this paper. From the 657 players considered, we used R to select four random players to see how our model both fits and adapts to the different careers of players. These players are: Robin Yount ([@sec-yount]), Pat Tabler ([@sec-tabler]), Junior Felix ([@sec-felix]), and Kendrys Morales([@sec-morales]).

@@

In explore results, discuss both the trajectories of about 4-5 players and show how they change (probably want to sample about 10 just to be safe). But then, just for fun, look at which player has the highest $\alpha$ level since that would be my models prediction of best HR hitter independent of everything else. It will probably be Barry Bonds.

@@

#This gives us a list of 10 players to consider so if there are a couple that look similar we can choose to not include them in the analysis
select_players <- Batting %>% 
  distinct(playerID) %>% 
  pull(playerID) %>% 
  sample(size=10, replace=FALSE)

select_players
```

@@

## General

This section should be dedicated to talking about general fit of the model. Explore using the normal plots that we've been using with the confidence intervals to see if the parameters are near their priors. And, also explore using the trace-plots. The linegraphs that go all over the place.

@@

In @fig-p651Pred we can examine how our model predicted Yount's HR probability, $\pi$, to change over his career. The x-axis is the same as it is in @fig-p651Act, where it describes Yount's deviation from the cetnered age (which is 30). The y-axis describes Yount's predicted HR probability $\pi$. An important note is that the y-axis is not measured on the logistic scale. We calculated it on the logistic scale but then converted it back to the normal scale, and use the normal scale in the figure, to prevent any confusion. Interpreting the points in @fig-p651Pred is very similar to the interpretation of points for @fig-p651Act. Each point in @fig-p651Pred represents the model's prediction of Yount's $\pi$ for a centered age value. For example, the point (0, 3.25) represents the model predicting that Yount's $\pi=3.25$ when he is 30 years old, which was very close to the actual value. Similarly, the points (-10, 1) and (6, 2) represent our model's prediction that Yount's $\pi=1$ when he is 20 years old and $\pi=2$ when he is 36 years old, respectively, which is also very close to what we saw in @fig-p651Act. Included in @fig-p651Pred is a red trend line for the same reasons mentioned in discussing @fig-p651Act. The red trend line represents the overall trend of the data in the scatter-plot. It is not nearly as necessary here, because the data is so clearly clustered in a parabolic pattern, but it is helpful in future plots. We can see in @fig-p651Pred that the predictions from our model follow closely to what we told it to expect. There is an increase in HR probability, $\pi$, for Yount from when he is 18 to when he is 32. This is quite close to what we see in the figure representing his actual HR probability. Then we see a sharp decline in predicted HR probability for Yount as he ages past 32 until he retires at the age of 38.

@@

See @fig-p582Act and @fig-p582Pred.

## Junior Felix {#sec-felix}

```{r}
# Junior Felix (177) Estimates cell

@@

```{r}
#Create the trajectory DF for player 177
i177.Trajectory <- Batting %>% 
  #Grab the player we're interested in
  filter(player_index == 177) %>% 
  #Only keep columns we're interested in
  select(playerID, player_index, yearID, year_index, Age, cAge, cAgeSq, AB, HR, p, park_index) %>%
  #Put the alpha, beta, and eta parameters in the DF
  mutate(alpha = alpha_i177 %>% pull(mean), .after = player_index) %>%
  mutate(beta = beta_i177 %>% pull(mean), .after = cAge) %>%
  mutate(eta = eta_i177 %>% pull(mean), .after = cAgeSq) %>%
  #Set-up to add "delta" to the table because 177 may have been traded
  right_join(delta_i177 %>% select(group, mean), by = join_by(park_index==group)) %>% 
  mutate(delta = mean, .after = park_index) %>% 
  select(!mean) %>% 
  #Set-up to add "xi" to the table because 177 may have been traded
  right_join(xi_i177 %>% select(group, mean), by = join_by(year_index==group)) %>% 
  mutate(xi = mean, .after = year_index) %>% 
  select(!mean) %>% 
  #Create the predictions for the player
  mutate(pred = alpha + (beta*cAge) + (eta*cAgeSq) + (delta) + (xi) , .after = p) %>% 
  mutate(inv_logit_pred = inv.logit(pred), .after=pred)
```

See @fig-p177Act and @fig-p177Pred.

Probably don't need to include this examination. Paper will probably be long enough and he's not that interesting.

@@

## Goodness of Fit - $\alpha$

# Conclusions and Directions for Future Research

@@

```{r}
#| label: fig-BattingCor
corrplot(cor(Batting_cor), type = "upper", order = "hclust", tl.col = "black", tl.srt = 45)
```

@@

```{r}
#| label: fig-p177Act
#| fig-cap: "Actual Trajectory for player #177 - Junior Felix"
ggplot(data = i177.Trajectory)+
  aes(x = cAge, y = p)+
  geom_point()+
  geom_smooth(se=FALSE)+
  scale_y_continuous(labels = label_percent())+
  labs(title = "Actual Trajectory for player #177 - Junior Felix",
       x = "Centered Age (Player Age - 30)",
       y = "HR's per AB",
       caption = "General trend of data in blue.")
```

```{r}
#| label: fig-p177Pred
#| fig-cap: "Predicted Career Trajectory for #177 - Junior Felix"
ggplot(data = i177.Trajectory)+
  aes(x = cAge, y=inv_logit_pred)+
  geom_point()+
  geom_smooth(color="red", se=FALSE)+
  scale_y_continuous(labels = label_percent())+
  labs(title = "Predicted Career Trajectory for #177 - Junior Felix",
       x = "Centered Age (Player Age - 30)",
       y = "HR's per AB",
       caption = "General trend of data in red.")
```

@@

```{r}
#| label: fig-alphaTrace
#| fig-cap: "Alpha Traceplots"
#651, 582, 177, 420
traceplot(fit.poly, pars = c("alpha[651]", "alpha[582]", "alpha[177]", "alpha[420]"))
```

>>>>

# Baseball Background {#sec-BaseBack}

@@

# Literature Review {#sec-LitReview}

@@

The goal of our model, as described in @eq-model, @eq-priors, and @eq-hyperpriors, was to be able to adaptively predict player $n$'s HR's. That is, we wished to create a model that would predict player $n$'s HR's and would adapt these predictions to be unique to player, $n$, the park played at, $p$, and the year played, $i$. To check the fit of their model, Fellingham and Fisher (2017) selected popular players (such as Hank Aaron and Barry Bonds) as a representative sample of the player population. However, the problem with this approach is that Fellingham and Fisher (2017) have selected players who are phenomenal hitters and generally outliers in the data. For example, Barry Bonds currently holds the single season and career HR records. Which means that he has the most HR's hit in one season and the most hit over the course of a career compared to all people who have every played in MLB. The career HR record he holds was previously held by Hank Aaron. Obviously, this sample of players is not representative of the population of players considered by Fellingham and Fisher (2017). So, while it may make the paper more enjoyable to read, it does not provide use with a rigorous understanding of the general fit of this model. We believe that a random sample should have been used to examine the fit of their model since it would give a better sense of overall model fit. Which, is exactly what we did in this paper. From the 657 players considered, we used R to select four random players to see how our model both fits and adapts to the different careers of players. These players are: Robin Yount ([@sec-yount]), Pat Tabler ([@sec-tabler]), and Kendrys Morales([@sec-morales]).

@@

@@

#This gives us a list of 10 players to consider so if there are a couple that look similar we can choose to not include them in the analysis
select_players <- Batting %>% 
  distinct(playerID) %>% 
  pull(playerID) %>% 
  sample(size=10, replace=FALSE)
```

@@

@@

In @fig-p651Pred we can examine how our model predicted Yount's HR probability, $\pi$, to change over his career. The x-axis is the same as it is in @fig-p651Act, where it describes Yount's deviation from the centered age (which is 30). The y-axis describes Yount's predicted HR probability $\pi$. An important note is that the y-axis is not measured on the logistic scale. We calculated it on the logistic scale but then converted it back to the normal scale, and use the normal scale in the figure, to prevent any confusion. Interpreting the points in @fig-p651Pred is very similar to the interpretation of points for @fig-p651Act. Each point in @fig-p651Pred represents the model's prediction of Yount's $\pi$ for a centered age value. For example, the point (0, 3.25) represents the model predicting that Yount's $\pi=3.25$ when he is 30 years old, which was very close to the actual value. Similarly, the points (-10, 1) and (6, 2) represent our model's prediction that Yount's $\pi=1$ when he is 20 years old and $\pi=2$ when he is 36 years old, respectively, which is also very close to what we saw in @fig-p651Act. Included in @fig-p651Pred is a red trend line for the same reasons mentioned in discussing @fig-p651Act. The red trend line represents the overall trend of the data in the scatter-plot. It is not nearly as necessary here, because the data is so clearly clustered in a parabolic pattern, but it is helpful in future plots. We can see in @fig-p651Pred that the predictions from our model follow closely to what we told it to expect. There is an increase in HR probability, $\pi$, for Yount from when he is 18 to when he is 32. This is quite close to what we see in the figure representing his actual HR probability. Then we see a sharp decline in predicted HR probability for Yount as he ages past 32 until he retires at the age of 38.

@@

Consider player $n=582$. That is, consider the player Pat Tabler. In @fig-p582Act we can see how Tabler's actual HR probability has changed over the years that he played. The x-axis of @fig-p582Act describes Tabler's deviation from the centered age, where the centered age value of 0 represents a player being 30 years old. And the y-axis represents the actual HR probability of Tabler. Thus, the point (-1, 2) tells us that Tabler's actual HR probability was 2% when he was 29 years old. In fact, the interpretation of @fig-p582Act is exactly the same as it was for @fig-p651Act, expect the probabilities are for Pat Tabler instead of Robin Yount. Observant readers may have noticed that there are two data points when $x=0$ and when $x=2$. This is because Pat Tabler was traded twice, once when he was 30 and another time when he was 32 years old. Since he was traded, it meant that he had two different home ballparks for that season. Thus, he had two different HR probabilities for that year with each representing his HR probability for a specific home ballpark. This "double observation" does not impact our impact our model at all. It is simply a quirk of the data. Included in @fig-p582Act is a blue trend line that snakes through the data. This line does not represent anything meaningful, it was included simply to make seeing and interpreting the trend present in the scatter-plot easier. To that end, we can see in @fig-p582Act that for the first 7 years of his career Tabler saw an increase in his actual HR probability, which lasted until he was 29 years old. Tabler's average HR probability at this time was about 1.5% during this period. Every year after he turned 29, with the exception of when he was 32, he performed starkly worse than he did in the beginning of his career. The average of these points being around 0.5%. The only outlier in this period was the season where he was 32, where he was traded to the New York Mets and had a higher HR probability for them that year then he did with the Kansas City Royals. However, this higher HR probability was probably due to sample size. He had the same amount of HR's with both teams but more than 4 times as many AB's with the Royals compared with the Mets. Overall, we can see that Tabler's career was different than Yount's in a couple key ways. The first is that Tabler stays well below the average HR probability for all of his career, whereas Yount says very close to it. The second way is that Tabler peaks before he is 30 and does so bad for the final years of his career that there is not much decline. Thus, it will be interesting to see how our adapts to this very interesting and unique player.

In @fig-p582Pred we can examine how our model predicted Tabler's HR probability, $\pi$, to change over his career. The x-axis is the same as it is in @fig-p582Act, where it describes Tabler's deviation from the centered age (which is 30). The y-axis describes Tabler's predicted HR probability $\pi$. An important note is that the y-axis is not measured on the logistic scale. We calculated it on the logistic scale but then converted it back to the normal scale, and use the normal scale in the figure, to prevent any confusion. Interpreting the points in @fig-p582Pred is very similar to the interpretation of points for @fig-p582Act. Each point in @fig-p582Pred represents the model's prediction of Tabler's $\pi$ for a centered age value. For example, the point (-1, 1.55) represents the model predicting that Tabler's $\pi=1.55$ when he is 29 years old. Similarly, the points (-4, 1.2) and (4, 0.8) represent our model's prediction that Tabler's $\pi=1.2$ when he is 26 years old and $\pi=0.8$ when he is 34 years old, respectively. However, the $\pi$ values were not nearly as accurate for Tabler as they were for Yount. We can see that the high variation of the actual HR probabilities has forced our model rely more on the priors than on the actual data. That is, we can see our estimates for $\pi$ are often much closer to the average HR rate across players, which is 3%, than they are to Tabler's specific data because Tabler is such an outlier. Included in @fig-p582Pred is a red trend line for the same reasons mentioned in discussing @fig-p582Act. The red trend line represents the overall trend of the data in the scatter-plot. Coming into this, the worry was that our model would rely too heavily on its priors and not be able to update its parabolic shape to match the strange career that Tabler had. However, after quickly examining the red line anyone can see that it has clearly updated to fit Tabler's data much better than @fig-p651Pred. That is, we can see that the red line from @fig-p582Pred matches the blue line in @fig-p582Act fairly well, demonstrating that the model updates itself based on the data it is fed. We can see that red line captures Tabler's beginning career well, with a slight increase in HR probability until he is 29. This is then followed by a steep and drastic drop in predicted HR rate for all years including and past when Tabler is 30 years old. Thus, although the actual proportions themselves do not line up as well as we may have hoped, we can see that the model does match trends well, which is a good sign. Thus, it will be interesting to see how different extreme player trajectories challenge the model.

```{r}
#| eval: false
# Junior Felix (177) Estimates cell

@@

```{r}
#| eval: false
#Create the trajectory DF for player 177
i177.Trajectory <- Batting %>% 
  #Grab the player we're interested in
  filter(player_index == 177) %>% 
  #Only keep columns we're interested in
  select(playerID, player_index, yearID, year_index, Age, cAge, cAgeSq, AB, HR, p, park_index) %>%
  #Put the alpha, beta, and eta parameters in the DF
  mutate(alpha = alpha_i177 %>% pull(mean), .after = player_index) %>%
  mutate(beta = beta_i177 %>% pull(mean), .after = cAge) %>%
  mutate(eta = eta_i177 %>% pull(mean), .after = cAgeSq) %>%
  #Set-up to add "delta" to the table because 177 may have been traded
  right_join(delta_i177 %>% select(group, mean), by = join_by(park_index==group)) %>% 
  mutate(delta = mean, .after = park_index) %>% 
  select(!mean) %>% 
  #Set-up to add "xi" to the table because 177 may have been traded
  right_join(xi_i177 %>% select(group, mean), by = join_by(year_index==group)) %>% 
  mutate(xi = mean, .after = year_index) %>% 
  select(!mean) %>% 
  #Create the predictions for the player
  mutate(pred = alpha + (beta*cAge) + (eta*cAgeSq) + (delta) + (xi) , .after = p) %>% 
  mutate(inv_logit_pred = inv.logit(pred), .after=pred)
```

@@

Consider player $n=420$. That is, consider the player Kendrys Morales. In @fig-p420Act we can see how Morales' actual HR probability has changed over the years that he played. The x-axis of @fig-p420Act describes Morales' deviation from the centered age, where the centered age value of 0 represents a player being 30 years old. And the y-axis represents the actual HR probability of Morales. Thus, the point (0, 4.25) tells us that Morales' actual HR probability was 4.25% when he was 30 years old. In fact, the interpretation of @fig-p420Act is exactly the same as it was for both @fig-p651Act and @fig-p582Act, expect the probabilities are for Kendrys Morales instead of Pat Tabler or Robin Yount. Observant readers may have noticed that there are two data points when $x=1$ and when $x=6$. This is because Morales was traded twice, once when he was 31 and another time when he was 36 years old. Since he was traded, it meant that he had two different home ballparks for that season. Thus, he had two different HR probabilities for that year with each representing his HR probability for a specific home ballpark. This "double observation" does not impact our impact our model at all. It is simply a quirk of the data. Included in @fig-p420Act is a blue trend line that snakes through the data. This line does not represent anything meaningful, it was included simply to make seeing and interpreting the trend present in the scatter-plot easier. To that end, we can see in @fig-p420Act that for the first four years of his career Morales saw and increase in his actual HR probability, which lasted until he was 27 years old. Over this period, his actual HR probability rose from 2.25% when he was a rookie, when he was 23 years old, to 6% when he was 27 years old. For the next four years of his career, there was a clear and sharp decline in his actual HR probability. His HR probability dropped from about 5.5%, when he was 28 years old, to 3.1% or 0.5% when he was 31 years old. Then, he saw an average increase in his HR probability to about 5% from when he was 32 years old until he was 35 years old. During the last year of his career, Morales was traded a final time and performed the worst he had done over his career. Between both teams he had a HR probability below 2%. Morales' career has been the most atypical compared to Yount and Tabler. We see him perform astonishingly well to begin and then perform much more poorly in the middle of his career when the average player excels. Then we see him make a recovery towards the endof his career, only for the last year of his career to be quite poor. This makes Morales' career very different than compared to the typical player, such as Yount, so this prediction for Morales will be one of the best tests we can provide to see how our model handles updating posteriors on non-standard data.

In @fig-p420Pred we can examine how our model predicted Morales' HR probability, $\pi$, to change over his career. The x-axis is the same as it is in @fig-p420Act, where it describes Morales' deviation from the centered age (which is 30). The y-axis describes Morales' predicted HR probability $\pi$. An important note is that the y-axis is not measured on the logistic scale. We calculated it on the logistic scale but then converted it back to the normal scale, and use the normal scale in the figure, to prevent any confusion. Interpreting the points in @fig-p420Pred is very similar to the interpretation of points for @fig-p420Act. Each point in @fig-p420Pred represents the model's prediction of Morales' $\pi$ for a centered age value. For example, the point (0, 4.25) represents the model predicting that Tabler's $\pi=4.25$ when he is 30 years old. Similarly, the points (-3, 4.75) and (4, 4.9) represent our model's prediction that Tabler's $\pi=4.75$ when he is 27 years old and $\pi=4.9$ when he is 34 years old, respectively. These $\pi$ values are fairly accurate, similar to what they were for Yount. However, similar to Tabler, we can see that for the seasons that Morales performed extraordinarily well or poorly, our model tends to fit more towards the average than how Morales actually performed. For example, when he was 27 years old Morales had an actual HR probability of 6%, yet our model only predicts for him to hit 4.75%. Similarly, when he was 36 years old Morales actually had a HR probability of less than 2%, yet our model predicts that he would have a HR probability of about 4%. This is likely because those extreme seasons were outliers both for Morales and across the data, so our model has a hard time "believing" that they are true more than they are a fluke. This is one of the properties of multilevel modeling, without a lot of evidence to the contrary it will assume that the player is more average than the observations suggests, which becomes helpful in the long run for prediction.

Included in @fig-p420Pred is a red trend line for the same reasons mentioned in discussing @fig-p420Act. The red trend line represents the overall trend of the data in the scatter-plot. Coming into this, the worry was that our model would rely too heavily on its priors and not be able to update its parabolic shape to match the strange career that Morales had. However, after quickly examining the red line anyone can see that it has clearly updated to fit Morales' strange career. That is, we can see that the red line from @fig-p420Pred matches the blue line in @fig-p420Act fairly well, demonstrating that the model updates itself based on the data it is fed. We can see that red line captures Morales' beginning career well, with an increase in HR probability until he is about 27 years old. This is then followed by a steep and drastic drop in predicted HR rate until he is about 31 years old. Then there is a recovery in $\pi$ until Morales is 34 where we again see a drop in $\pi$ as Morales finishes out his career. Thus, we can see that the model does match trends well. It captures both the peaks that we saw for Morales throughout his career and matches them up to the correct time and gets rougly the same HR probabilities that we saw in the actual data. Therefore, we can be confident in saying that for Morales, our model updates its expectations based on the data well.

## Goodness of Fit - $\alpha$ {#sec-alpha}

There are many numerical measures of goodness-of-fit statistics because researchers are often searching for ways to determine which models of their many mondels are the best for their particular goals. Thus, many of these numerical estimates should be used in a comparative sense, with and against models which are trained on the same data set and attempt to answer the same resarch question. Examples of these statistics are the Bayesian Information Criterion (BIC), the Widely Applicable Information Criterion (WAIC), the Akaike Information Criterion (AIC), and many others. While it would be possible to quantify these statistics for our model, it would be frivolous. Due to computational constraints, we were only able to generate one model to answer our research question, and it has been the one discussed extensively in this paper. Hence, generating these goodness-of-fit statistics would be useless because they should be used ina comparative context. But, there are no models with which to compare our model to. Our model uses unique data compared to the models which attempt to answer the same research question. Thus, comparing models with these statistics would be foolish.

However, that does not mean that there are no ways to check how well the model has fit. Beyond what has been discussed in @sec-yount, @sec-tabler, and @sec-morales, there is another method to check the goodness-of-fit for a model. Trace-plots (sometimes called trank-plots) are a way to check how individual parameters have fit. We have examined how the $\alpha$ parameter has fit for four players: Yount (651), Tabler (582), Morales (420), and Junior Felix (177). Junior Felix has not been discussed yet because he was the fourth of the random player's sampled and did not have as interesting of a career as the other three players discussed. Since he did not have a very interesting career, he did not provide much insight into the model fit and updating, so we decided to not display his career trajectory. However, we have included it @fig-alphaTrace as the last graph in the bottom right because it is a good example of what a trace-plot should look like, which contrasts nicely with the other trace-plots.

First examine the trace-plot, in the bottom right of @fig-alphaTrace, labeled "alpha\[177\]". This plot represents our models Markov Chain Monte Carlo (MCMC) fitting process. What we should see, if the fit went well, is that the chains (the different colored lines) should remain at roughly the same place over the duration of the iterations. This is what we see in "alpha\[177\]". We can see that, in general, the chains stay between -3.5 and -3 over their 5,000 iterations. There are no sudden deviations so we can conclude that "alpha\[177\]" or $\alpha_{177}$ in our model, has a good fit. This stands in contrast to the other fits for $\alpha$. For example, "alpha\[651\]" and "alpha\[420\]" both see a slight upward trend in their $\alpha$ estimates from iterations 5,000 to 7,000. Then after about iteration 7,000 we see a sharp decrease in both estimates. For "alpha\[651\]" this drop goes from about -3.1 to about -3.3. While for "alpha\[420\]" this drop goes from about -2.75 to -3. Furthermore, we can see fluctuation in both "alpha\[651\]" and "alpha\[420\]" from iterations 7,000 to 10,000 suggesting that even after the shift the $\alpha$ parameter still does not find a good fit. The same problems persist for "alpha\[582\]" but are not as pronounced. At around iteration 7,000 our average for "alpha\[582\]" drops from about -4.1 to -4.3. The fluctuations for the average value of "alpha\[582\]" are not as pronounced as they were for either "alpha\[651\]" or "alpha\[420\]", and actually look quite similar to the estimates for "alpha\[177\]".

Overall, we can see that the fit for our $\alpha$ parameter is quite poor for the consdiered players, even though the our model still fits the data well. This suggests that the other parameter fit well since our model did a good job at predicting and adapting to players trajectories. Moreover, in the future if we wish to get a better fit we should consider running our MCMC chain for more iterations. Running for more iterations should allow the model to find the best parameter value, which is what the trace-plots visualize. This is likely why Fellingham and Fisher (2017) ran their model for 200,000 iterations. Allowing the model to iterate for that long will let it find and settle on a parameter value.

## Goodness of Fit - $\beta$ and $\eta$ {#sec-betaandeta}

We can also examine the parameter fits for both $\beta_n$ and $\eta_n$, for $n\in\{651,582,420,177\}$. We chose to group $\beta$ and $\eta$ together in this section because in general they capture the same effect. That effect being how aging impacts player $n$'s ability to hit HR's. Furthermore, one can easily tell from @fig-betaTrace and @fig-etaTrace that the trace-plots for $\beta$ and $\eta$ are not nearly as interesting as they were for $\alpha$ in @sec-alpha.

Consider first the fits for $\beta$ in @fig-betaTrace. The fits presented are for Robin Yount (651), Pat Tabler (582), Kendrys Morales (420) and Junior Felix (177). All fits in this figure look exactly as a good trace-plot should. There are no sudden shifts in the model's estimates for $\beta_n$. Nor are there any serious fluctuations throughout the 5,000 to 10,000 iterations. They all stay at roughly the same value throughout the iterations. We can tell this is not due to scaling on the y-axis because, compared to @fig-alphaTrace, the y-axis of @fig-betaTrace is much more concentrated and has a range of no more that 0.2. Thus, since all the chains stay at roughly the same value over the iterations we know that our model has a good fit for $\beta$ for our four players considered.

Consider second the fits for $\eta$ in @fig-etaTrace. The discussion of these fits will be almost identical to the fits for $\beta$ because the model also fit very well here for the four players considered. The fits presented are for Robin Yount (651), Pat Tabler (582), Kendrys Morales (420) and Junior Felix (177). All fits in this figure look exactly as a good trace-plot should. There are no sudden shifts in the model's estimates for $\eta_n$. Nor are there any serious fluctuations throughout the 5,000 to 10,000 iterations. They all stay at roughly the same value throughout the iterations. We can tell this is not due to scaling on the y-axis because, compared to @fig-alphaTrace, the y-axis of @fig-etaTrace is much more concentrated and has a range of no more that 0.04. Thus, since all the chains stay at roughly the same value over the iterations we know that our model has a good fit for $\eta$ for our four players considered.

Most other Bayesian papers would continue this discussion for their other parameters, which for our model would be discussing $\delta$ and $\xi$. However, these parameters in our model are not player specific, which means to consider them would mean examining 14 $\delta$'s and 33 $\xi$'s. Discussion of of 47 additional parameters in the context of each player seems both trivial and cumbersome. Examining these parameters in the context of the player would not improve our understanding of the model fit. In fact, after examining these parameters we would simply be left with how our model got its estimates which we have already reviewed in @sec-yount, @sec-tabler, and @sec-morales. Thus, because the labor is high and the pay-off is low we have elected to let the reader review any additional goodness-of-fit figures and statistics they would like. The code that we used to generate the model is available on [GitHub](https://github.com/Sturner21/MATH-470) and readers can use it to replicate our findings and investigate our model for themselves.

# Conclusions and Directions for Future Research {#sec-conclusion}

Overall, our model for predict the HR production of MLB hitters appears to do a fair job. It does a good job at updating its predictions based on the data for each player. But, still keeps its predictions close to the mean unless there is strong evidence, like with Pat Tabler, to suggest that it should deviate. Furthermore, as we can see with the contrast between Robin Yount and Kendrys Morales, our model does not have to rely solely on what the priors tell it. Our model can update itself to either follow a typical career trajectory, as we saw with Robin Yount, and it can update itself to follow an atypical career trajectory, as we saw with Kendrys Morales. Thus, we can conclude that our model does a good job at fitting overall. However, this model is far from perfect. We could see with predictions from Pat Tabler that our model would not predict HR probabilities $\pi$ at a rate where gambling would be a wise endeavor.

@@

```{r}
#| label: fig-BattingCor
#| fig-cap: "Batting Correlation Table"
corrplot(cor(Batting_cor), type = "upper", order = "hclust", tl.col = "black", tl.srt = 45)
```

@@

```{r}
#| label: fig-p177Act
#| fig-cap: "Actual Trajectory for player #177 - Junior Felix"
#| eval: false
ggplot(data = i177.Trajectory)+
  aes(x = cAge, y = p)+
  geom_point()+
  geom_smooth(se=FALSE)+
  scale_y_continuous(labels = label_percent())+
  labs(title = "Actual Trajectory for player #177 - Junior Felix",
       x = "Centered Age (Player Age - 30)",
       y = "HR's per AB",
       caption = "General trend of data in blue.")
```

```{r}
#| label: fig-p177Pred
#| fig-cap: "Predicted Career Trajectory for #177 - Junior Felix"
#| eval: false
ggplot(data = i177.Trajectory)+
  aes(x = cAge, y=inv_logit_pred)+
  geom_point()+
  geom_smooth(color="red", se=FALSE)+
  scale_y_continuous(labels = label_percent())+
  labs(title = "Predicted Career Trajectory for #177 - Junior Felix",
       x = "Centered Age (Player Age - 30)",
       y = "HR's per AB",
       caption = "General trend of data in red.")
```

@@

```{r}
#| label: fig-alphaTrace
#| fig-cap: "Alpha Traceplots"
#651, 582, 177, 420
traceplot(fit.poly, pars = c("alpha[651]", "alpha[582]", "alpha[420]", "alpha[177]"))
```

```{r}
#| label: fig-betaTrace
#| fig-cap: "Beta Traceplots"
traceplot(fit.poly, pars = c("beta[651]", "beta[582]", "beta[420]", "beta[177]"))
```

```{r}
#| label: fig-etaTrace
#| fig-cap: "Eta Traceplots"
traceplot(fit.poly, pars = c("eta[651]", "eta[582]", "eta[420]", "eta[177]"))
```
